---
title: "Ant_Bird_Full_Analysis"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(MASS)
library(pscl)
library(multcomp)
library(emmeans)
library(broom)
```

```{r}
# Data cleaning and transformation steps
# For example:
# bird_ind <- bird %>%
#   mutate(
#        rsa = (Rinter > 0),
#        roa = (Rintra > 0),
#        isa = (Iinter > 0),
#        ioa = (Iintra > 0),
#     )
# write_csv(bird_ind, "bird_ind.csv")

# apply factor
hierarchy_fct <- c("Phleg", "M. fort", "Rheg", "D. mer", "G. sal") # most dominant to least dominant

# load in dataset
bird_ind <- read_csv("data/bird_ind.csv") %>% mutate(species = factor(species, hierarchy_fct))
head(bird_ind)
# verify factor levels
levels(bird_ind$species)
```

# Model Selection

## Best model selection for Successes as Response

We have nbinom, poisson, zero-infl and quasi forms of these models, as possible candidates

### Fit Poisson model

```{r}
pois_full <- glm(Successes ~ species + rsa + roa + isa + ioa + species:rsa + species:roa + species:isa + species:ioa, family = "poisson", data = bird_ind)
summary(pois_full)
```

### Fit Quasi Poisson

```{r}
quasi_full <- glm(Successes ~ species + rsa + roa + isa + ioa + species*rsa + species*roa + species*isa + species*ioa, family = "quasipoisson", data = bird_ind)
summary(quasi_full)
```

The dispersion parameter is not much larger than 1 meaning there isn't much overdispersion to account for. For example, if we multiply each of the standard errors of the original poisson model to the sqrt(dispersion parameter), there would not be much of a difference between the two model types.

### Fit Negative Binomial model

```{r}
nbinom_full <- glm.nb(Successes ~ species + rsa + roa + isa + ioa + species:rsa + species:roa + species:isa + species:ioa, data = bird_ind) summary(nbinom_full)
```


Ruled out Zero-Inflated Negative Binomial and Poisson models due to species:ioa and species:isa having NA values for G. Sal and M. Fort. 

### Negative Binomial or Poisson?

```{r}
bird_ind %>%
  group_by(species) %>%
  summarize(
    meanRinter = mean(Rinter),
    meanRintra = mean(Rintra),
    meanIinter = mean(Iinter),
    meanIntra = mean(Iintra),
    sdRinter = sd(Rinter),
    sdRintra = sd(Rintra),
    sdIinter = sd(Iinter),
    sdIintra = sd(Iintra),
    n = n()
  )
```
Upon looking at the shape of data, it is revealed that most standard deviation counts are higher than the predictor means suggesting our mean and variances are not equal. Therefore, Poisson is not appropriate for our data. However, negative binomial is able to account for overdispersion in cases where the variance is greater than the mean. 

### Fitting Best Model: Testing Predictor Significance

```{r}
nbinom_full <- glm.nb(Successes ~ species + rsa + roa + isa + ioa + species*rsa + species*roa + species*isa + species*ioa, data = bird_ind)
summary(nbinom_full)
plot(nbinom_full)
```

```{r}
# Use stepwise function to determine model with lowest AIC
step.model <- step(nbinom_full, direction="both")
```

```{r}
# best model candidates w/ lowest AICs:
bestmod1 <- glm.nb(Successes ~ species + rsa + ioa + species*rsa, data = bird_ind)
bestmod2 <- glm.nb(Successes ~ species + rsa + species*rsa, data = bird_ind)
summary(bestmod1)
summary(bestmod2)
```

Since there is not a large difference in AIC between the two candidates, we choose the simpler option: bestmod2.

```{r}
# assess goodness of fit for model
gof.pvalue = 1 - pchisq(bestmod2$deviance, bestmod2$df.residual)
gof.pvalue
```
We have a very low p-value meaning our model doesn't fit the data very well. This means we havve a 2% chance of observing a test statistic as, or more extreme than, the one calculated from the data under the null hypothessis. 

## Best Model Selection for Attempts as Response
```{r}
nbinom_full_a <- glm.nb(attempts ~ species + rsa + roa + isa + ioa + species*rsa + species*roa + species*isa + species*ioa, data = bird_ind)
summary(nbinom_full_a)

# use stepwise function to determine model with lowest AIC
step.model <- step(nbinom_full_a, direction="both")
```
```{r}
# best model candidates w/ lowest AICs:
bestmod1_a <- glm.nb(attempts ~ species + roa + ioa + species*roa, data = bird_ind)
bestmod2_a <- glm.nb(attempts ~ species + rsa + roa + isa + ioa + species:roa, data = bird_ind)
summary(bestmod1_a)
summary(bestmod2_a)
```

```{r}
anova(bestmod1_a, bestmod2_a, test = "Chisq")

# Since there's no siginificant difference between the two models' AIcs, and 
# there is no evidence to reject that the models are any different, we're going to chose the first model.

bestmod_a <- bestmod1_a
```

```{r}
# assess goodness of fit for model
gof.pvalue = 1 - pchisq(bestmod_a$deviance, bestmod_a$df.residual)
gof.pvalue
# low p-value; model does not fit the data very well...
```


# Best Model Interpretations

## Pairwise Comparisons for Successes
Performed pairwise comparisons across the significant interaction term of rsa found in the best model. 

```{r}
# visualize the log mean number of successes (linear prediction) for species receiving same species aggression (rsa):
emmip_successes <- emmip(bestmod, species ~ rsa | rsa) + labs(y = "Log Mean Number of Successes", x = "Levels of Rsa")

# ggsave("data/log_mean_rsa.jpg", emmip_successes, width = 8, height = 6, units = "in")
```

```{r}
# calculate estimated marginal means (EMMs) for pairwise comparisons between species w respect to rsa (grouping var)
emmeans(bestmod, pairwise ~ species | rsa)
```

```{r}
# calculate marginal means for the interaction between species and rsa
bird.emm <- emmeans(bestmod, ~ species * rsa)
# create pairwise contrasts between rsa levels, adjust for multiple testing using the multivariate t method
bird.cont <- contrast(bird.emm, "consec", simple = "each", combine = TRUE, adjust = "mvt")

# Estimate marginal means for the interaction
emm_int <- emmeans(bestmod, specs = ~ species * rsa)

# Pairwise comparisons with adjustment for multiple testing
comp <- contrast(emm_int, method = "pairwise", adjust = "bonferroni")
summary(comp)
```

```{r}
confint_successes <- 
  plot(comp) + theme_bw() + geom_vline(xintercept = 0, color = "red") + 
  labs(y = "Estimated Log-Count Difference", x = "95% Confint Estimate")
confint_successes
ggsave("data/confint_successes.jpg", plot = confint_successes, width = 7, height = 10, dpi = 300)

# visualization of confidence intervals; any interval that contains 0 is not significant 
plot(comp)
```

## Pairwise Comparisons for Attempts
Performed pairwise comparisons across the significant interaction term of roa found in the best model.
```{r}
# visualize the log mean number of successes (linear prediction) for species receiving other species aggression (roa)
emmip_attempts <- emmip(bestmod_a, species ~ roa | roa) + labs(y = "Log Mean Number of Attempts", x = "Levels of Roa" )

# ggsave("data/log_mean_roa.jpg", emmip_attempts, width = 8, height = 6, units = "in")
```

```{r}
# calculate estimated marginal means (EMMs) for pairwise comparisons between species w respect to roa (grouping var)
emmeans(bestmod_a, pairwise ~ species | roa)
```

```{r}
# calculate marginal means for the interaction between species and roa
bird.emm <- emmeans(bestmod_a, ~ species * roa)
# create pairwise contrasts between roa levels, adjust for multiple testing using the multivariate t method
bird.cont <- contrast(bird.emm, "consec", simple = "each", combine = TRUE, adjust = "mvt")

# Estimate marginal means for the interaction
emm_int <- emmeans(bestmod_a, specs = ~ species * roa)

# Pairwise comparisons with adjustment for multiple testing
comp <- contrast(emm_int, method = "pairwise", adjust = "bonferroni")
summary(comp)
```
```{r}
# visualization of confidence intervals; any interval that contains 0 is not significant 
confint_attempts <- 
  plot(comp) + theme_bw() + geom_vline(xintercept = 0, color = "red") + 
  labs(y = "Estimated Log-Count Difference", x = "95% Confint Estimate")
confint_attempts
# ggsave("data/confint_attempts.jpg", plot = confint_attempts, width = 7, height = 10, dpi = 300)

# calculate estimated marginal means (EMMs) for pairwise comparisons between species w respect to roa (grouping var)
emmeans(bestmod_a, pairwise ~ species | roa)
```














